The Concurrent Calculi Formalisation Benchmark

### 2023-10-05 paper/paper.tex
Minor fixes in introduction.

The introduction of the influential POPLMark challenge~\cite{POPLMark}
spearheaded a shift to including mechanised proofs accompanying
published papers. POPLMark's abstract starts with an important
question ``How close are we to a world where every paper on program-
ming languages is accompanied by an electronic appendix with machine-
checked proofs?''. Much progress has been done in that respect,
notably the POPLMark reloaded~\cite{POPLMarkReloaded} challenge aims
to identify and develop best practices and tool support for proofs
using logical relations.

The authors of both POPLMark and POPLMark Reloaded
were careful to note that their benchmarks were only a beginning, and
they specifically mention reasoning about concurrency using
coinduction and linear environments as points for future work. Papers
in concurrency theory do often include mechanised proofs
(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also
directly discuss formalisation techniques
(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}). However,
our experience is that choosing techniques and tools remains a
challenge with few guidelines. We propose a new collection of
benchmark problems specifically designed to tackle the issues
encountered when mechanising message-passing concurrency using process
calculi. A goal of our collection of benchmark problems is to push for
the development of guidelines and tutorials, and our challenge
problems are a simple setting in which to demonstrate and compare
techniques. Another goal is to strengthen the culture of mechanisation
by rallying the community to collaborate on exploring and developing
new tools and techniques. We have identified three key aspects which
cause issues when mechanising concurrency theory: linearity, scope
extrusion in name passing systems and infinite behaviour modelled
coinductively. Each part of our benchmark exercises the three aspects
independently to enable comparison and to keep the challenge problems
small. The three aspects are of course not the only ones that may
cause issues in mechanisations, but they are fundamental to the field
and will thus appear in most mechanisations of concurrency theory.
Actual mechanisations will additionally often require more than just
one aspect, but we do not address the combination of techniques. In
the rest of this paper we discuss the design of the challenge and
briefly describe the three challenge problems.

>>>>

The introduction of the influential POPLMark challenge~\cite{POPLMark}
spearheaded a shift to including mechanised proofs accompanying
published papers. POPLMark's abstract starts with an important
question: ``How close are we to a world where every paper on program-
ming languages is accompanied by an electronic appendix with machine-
checked proofs?''. Much progress has been done in that respect,
notably the POPLMark Reloaded~\cite{POPLMarkReloaded} challenge aims
to identify and develop best practices and tool support for proofs
using logical relations.

The authors of both POPLMark and POPLMark Reloaded
were careful to note that their benchmarks were only a beginning, and
they specifically mention reasoning about concurrency using
coinduction and linear environments as points for future work. Papers
in concurrency theory do often include mechanised proofs
(e.g.~\cite{DBLP:conf/pldi/Castro-Perez0GY21,DBLP:conf/tacas/CastroFY20,lmcs:9985,
  DBLP:journals/jar/CruzFilipeMP23, Tirore:2023}), and others also
directly discuss formalisation techniques
(e.g.~\cite{DBLP:journals/jar/BengtsonPW16, DBLP:conf/tphol/Gay01,
  DBLP:conf/ppdp/Thiemann19, DBLP:conf/forte/ZalakainD21}). However,
our experience is that choosing techniques and tools remains a
challenge with few guidelines. We propose a new collection of
benchmark problems specifically designed to tackle the issues
encountered when mechanising message-passing concurrency using process
calculi. A goal of our collection of benchmark problems is to push for
the development of guidelines and tutorials, and our challenge
problems are a simple setting in which to demonstrate and compare
techniques. Another goal is to strengthen the culture of mechanisation
by rallying the community to collaborate on exploring and developing
new tools and techniques. We have identified three key aspects which
cause issues when mechanising concurrency theory: linearity, scope
extrusion in name passing systems and infinite behaviour modelled
coinductively. Each part of our benchmark exercises one of the three aspects
independently to enable comparison and to keep the challenge problems
small. The three aspects are of course not the only ones that may
cause issues in mechanisations, but they are fundamental to the field
and will thus appear in most mechanisations of concurrency theory.
Mechanisations of actual research will additionally often require more than just
one aspect, but we do not address the combination of techniques. In
the rest of this paper we discuss the design of the challenge and
briefly describe the three challenge problems.
